---
output: pdf_document
---

```{r setoptions,echo=TRUE,include=FALSE}
library(datasets); library(knitr); library(tidyverse); library(xtable); library(ggplot2)
library(grid); library(gridExtra)
options(xtable.comment = FALSE)
```

```{r computetime,echo=TRUE}
date <- format(Sys.time(), "%b %d %Y")
rand <- rnorm(1)
```

## Project 1: Daily Step Movement
Submitted by Andrew Chang on `r date` for the Johns Hopkins Reproducible Research course

```{r readdata,echo=TRUE}
readAct <- read.csv("C:/Users/arkai/Documents/R/projects/DSS05_RepRes/repdata_data_activity/activity.csv", 
                    colClasses = c("numeric", "POSIXct", "integer")) 
```

```{r daysteps,warning=FALSE,message=FALSE,fig.height=3,fig.width=4}
actDaily <- readAct %>%
        select(
                date, steps
        ) %>%
        group_by(date) %>%
        summarize(
                day_steps = sum(steps, na.rm = TRUE)
        )

actDailyMean <- round(mean(actDaily$day_steps, na.rm = TRUE), 1)
actDailyMed <- round(median(actDaily$day_steps, na.rm = TRUE), 1)

ggplot(data = actDaily, aes(day_steps)) + 
        geom_histogram()
```

The mean and median number of steps taken each day are `r actDailyMean` and `r actDailyMed`, respectively. 

```{r intvl,echo=TRUE,warning=FALSE,message=FALSE,fig.height=3,fig.width=4}
actIntvl <- readAct %>%
        select(
                date, steps, interval
        ) %>%
        group_by(interval) %>%
        summarize(
                avg_intvl_steps = mean(steps, na.rm = TRUE)
        )

maxIntvlSteps <- round(max(actIntvl$avg_intvl_steps, na.rm = TRUE), 1)
maxIntvl <- actIntvl$interval[which.max(actIntvl$avg_intvl_steps)]

ggplot(data = actIntvl, aes(interval, avg_intvl_steps)) + 
        geom_line()

nullStep <- readAct %>%
        select(
                steps, date, interval
        ) %>%
        filter(is.na(steps))

nullStepRows <- nullStep %>%
        count()

nullStepCount <- nullStepRows$n
```

The interval with the highest average daily step count is `r maxIntvl` with `r maxIntvlSteps` steps.

After the imputation of missing values, there are `r nullStepCount` missing values in the dataset.  


```{r updatemissing,echo=TRUE,warning=FALSE,message=FALSE,fig.height=3,fig.width=4}

avgSteps <- nullStep %>%
        inner_join(actIntvl, by = "interval") %>%
        select(
                date,
                interval,
                avg_intvl_steps
        )

readAct0 <- readAct %>%
        left_join(avgSteps, by = c("date", "interval")) %>%
        mutate(
                steps = ifelse(is.na(steps), avg_intvl_steps, steps)
        )
        
nullRA0 <- readAct0 %>%
        select(
                steps, date, interval
        ) %>%
        filter(is.na(steps))

actDaily_impAvg <- readAct0 %>%
        select(
                date, steps
        ) %>%
        group_by(date) %>%
        summarize(
                day_steps = sum(steps, na.rm = TRUE)
        )

ggplot(data = actDaily_impAvg, aes(day_steps)) + 
        geom_histogram()

```

```{r weekend,warning=FALSE,message=FALSE,fig.height=3,fig.width=4}

readAct_wkpt <- readAct0 %>%
        mutate(
                weekpart = factor(ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
        )

actDaily_wkptAvg <- readAct_wkpt %>%
        select(
                date, weekpart, steps
        ) %>%
        group_by(date, weekpart) %>%
        summarize(
                day_steps = sum(steps, na.rm = TRUE)
        )

ggplot(data = actDaily_wkptAvg, aes(day_steps)) + 
        geom_histogram() + 
        facet_grid(rows = vars(weekpart))

```

